---
title: "MostPopularChidrensNamesCleaning"
author: "Kay Ayala"
date: "12/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Library 


```{r library, echo=TRUE}
library(dplyr)
```

## Import Data from txt

```{r import, echo=TRUE}
names_2015 = read.table("~/Desktop/Doing_Data_Science/Homework_5/yob2015.txt", sep=",")
colnames(names_2015) = c("Name","Sex","amount_of_children") #renames columns in 2015 data
names_2016 = read.table("~/Desktop/Doing_Data_Science/Homework_5/yob2016.txt", sep=";")
colnames(names_2016) = c("Name","Sex","amount_of_children") #renames columns in 2016 data
```

## Prints Data Structure and Summary

``` {r prints_data_summary, echo = TRUE}
print(head(names_2016)) 
print(str(names_2016)) 
print(summary(names_2016))

```


## Removes Duplicates, Errors, and Displays last ten rows

```{r removes_duplicates, echo=TRUE}
names_2016 = names_2016[!duplicated(names_2016$Name),] # removes duplicates
names_2015 = names_2015[!duplicated(names_2015$Name),]

print(grep("yyy$", names_2016$Name))
y2016 = names_2016[-212,] # removes line with "fionayyy"
y2015 = names_2015 #renames dataset

print(y2015[(nrow(y2015)-10):nrow(y2015),]) # Displays last 10 rows in 2015 Data 
```

These ten rows are interesting because they are all fairly uncommon boy names that start with 'Z'. They also look like they are from languages other than English. 

## Merges Data Frames, Renames Columns
```{r merge_and_rename, echo=TRUE}
final = merge(y2015, y2016, by="Name", all=TRUE)

colnames(final) = c("Name", "Sex_2015", "amount_of_children_2015", 
                    "Sex_2016", "amount_of_children_2016")
```

##Removes Missing Values, Creates New Column 

New column containts total number of children per name in both years. 

```{r removes_missing_values, echo=TRUE}
final = final[!is.na(final$amount_of_children_2015),]
final = final[!is.na(final$amount_of_children_2016),]

#print(str(final))
final$amount_of_children_2015 = as.numeric(final$amount_of_children_2015)
final$amount_of_children_2016= as.numeric(final$amount_of_children_2016)

final$Total_number_of_children = final$amount_of_children_2015 + final$amount_of_children_2016
```

## Prints Dimensions

```{r prints_dimensions, echo=TRUE}
print(dim(final))
```

## Sorts Data Frame and Prints Top Ten Names

```{r sorts_and_prints, echo=TRUE}
final = final[order(final$Total_number_of_children, decreasing = TRUE),]

final = final[!final$Sex_2015=="M",]

cat("Top Ten Names\n")
print(head(as.character(final$Name),10))
```

## Removes Extranious Data Frame Values

Retains only top ten names and the total number of children who have them. 

```{r removes_extranious_data, echo=TRUE}
new_final = data.frame(final$Name, final$Total_number_of_children)
final=new_final
colnames(final) = c("Name", "Total_number_of_children")
```

## Writes Data Frame to CSV

```{r writes_data, echo=TRUE}

write.csv(file = "~/Desktop/Doing_Data_Science/Homework_5/MostPopularGirlsNames2015and2016.csv", x = final)

```

# Github Link 
Note: This file can be found at https://github.com/KayMSDS/PopularChildrenNames.git along with the data used in and produced by this assignment.  